#R markdown for Reproducible research Week 2 Assignment
======================================================

Loading and preprocessing the data
```{r}
setwd("C:/Users/Teng Family/Desktop/Ziyang/Data Science/MOOC courses/JHU data science specialization/Reproducible Research/Week 2")

dta<-read.csv("activity.csv")
dta$date<- as.Date(dta$date)
```

Calculating the total number of steps and Plotting a histogram for each day

```{r}
#Total number of steps for each day
total.steps<- tapply(dta$steps, dta$date, sum, na.rm= TRUE)

library("ggplot2")

#Plotting histigrom for the number of steps taken for each day
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
 

#Mean and median of total steps each day
mean(total.steps, na.rm = TRUE)
median(total.steps, na.rm = TRUE)
```

What is the average activity pattern
```{r}
average<- aggregate(x=list(steps= dta$steps), by=list(interval= dta$interval), mean, na.rm= TRUE)

ggplot(average, aes(x= interval, y= steps)) + geom_line() + xlab("5 minute intervals")+ylab("Average number of steps")
```

which 5-minute interval, on average across all the days in the dataset, which contains the maximum number of steps

```{r}
average[which.max(average$x),]
```

Imputing missing values

```{r}
#Calculate and report the total number of missing values in the dataset
sum(is.na(dta$steps))

#Using mean to replace the NA values
fill.values<- function(steps, interval){
  filled<- NA
  if (!is.na(steps))
      filled<- c(steps)
  else
      filled<- (average[average$interval==interval,"steps"])
  
}
filled.dta<- dta
filled.dta$steps<- mapply(fill.values, filled.dta$steps, filled.dta$interval)
```

Make a histogram of the total number of steps taken each day
```{r}
total_steps<- tapply(filled.dta$steps, filled.dta$date,sum)
qplot(total_steps, binwidth=1000, xlab= "total number of steps each day")
```

Mean and median of newly constructed data
```{r}
mean(total_steps)
median(total_steps)
```

Are there differences in activity patterns between weekdays and weekends?
```{r}
weekday.or.weekend<- function(date){
    days<-weekdays(date)
    if (days%in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
      return("weekday")
    else if (days %in% c("Saturday", "Sunday"))
      return("weekend")
    else
      stop("invalid date")
}
new.dta<- filled.dta
new.dta$day<- sapply(new.dta$date, weekday.or.weekend) 

#Making a panel plot of time series for average number of steps taken across weekdays or weekends
average.wkday.wkend<- aggregate(steps ~ interval + day, data=new.dta, mean)
ggplot(average.wkday.wkend, aes(interval, steps))+ geom_line()+ facet_grid(day ~ .) + xlab("5 minute interval") + ylab("Number of steps")
```

Activities seem to differ for weekdays and weekend separately